Observing Economic Growth in Unrecognized States with Nighttime Light \thanks{I thank Elizaveta Kugaevskaia for her superb research assistance on this project. Financial support from Japan Society for the Promotion of Science (JSPS Kakenhi Grant Number 17K13730) is gratefully acknowledged.}

### 2019-03-20 paper/paper.tex
Revise the introduction after proof-reading #54

\begin{abstract}
This paper uses the satellite images of nighttime light to estimate economic growth rates in four unrecognized states of the former Soviet Union: Nagorno-Karabakh in Azerbaijan, Abkhazia and South Ossetia in Georgia, and Transnistria in Moldova. We then compare these estimates against those similarly obtained for parent states to gauge the impact of non-recognition as sovereign states on economic activities. The estimated economic growth rates do not differ much between the breakaway territories and their parent states, suggesting that the economic impact of non-recognition as states may be very limited.
\end{abstract}

@@

\section{Introduction}
% Introducing frozen conflicts
Back in 1991, the Soviet Union broke up into fifteen of its constituent republics. 
Immediately after, wars of secession broke out in three of these newly independent states: Azerbaijan, Georgia, and Moldova.
Secessionists, consisting of minority ethnic groups in each republic, succeeded in taking over their territories and declared their independence from their parent states. 
These breakaway territories include Nagorno-Karabakh in Azerbaijan, Abkhazia and South Ossetia in Georgia, and Transnistria in Moldova (Figure \ref{map}). 
However, not a single country around the world recognized them as sovereign states. 
These four ``states'' remain internationally unrecognized until today.%
\footnote{%
	Since 2008, Russia and a few other countries have recognized Abkhazia and South Ossetia as sovereign states, though.
}

@@

The empirical challenge for such endeavor is the lack of data. International organizations such as the World Bank do not provide any data for unrecognized states. 
The limited state capacity of these territories implies that there is no official statistics. 
Any surveys conducted in their parent countries exclude these territories from the sampling procedure.%
\footnote{%
	As an example, the Demographic and Health Survey conducted in Moldova in 2005 excludes the entire Transnistria from its sample (National Scientific and Applied Center for Preventive Medicine 2006:\ xxvi).
} 
To overcome this data limitation, the present study uses the satellite images of nighttime light for the period 1993--2013 to recover the unobserved annual economic growth rates of the four unrecognized states in the former Soviet Union.

Specifically, we first estimate the elasticity of real GDP with respect to nighttime light intensity, conditional on country and year fixed effects, from the annual panel data of 177 countries around the world. 
Using this elasticity estimate and the estimated year fixed effects, we then recover the annual growth rates of real GDP for unrecognized states from their nighttime light intensity. 
We finally compare these recovered real GDP growth rates to the ones similarly estimated for parent countries. 
Under the assumption that parent countries provide a valid counterfactual, the difference in the annual growth rates can be interpreted as the impact of being unrecognized as sovereign states.

We find that the impact of unrecognized states is very limited in all of the four cases. 
The recovered annual economic growth rates do not differ much between the breakaway territories and their parent states.

@@

Second, the issue of unrecognized states has received no attention in the emerging literature on the impacts of conflicts.%
\footnote{
	See Blattman and Miguel (2010:\ Section 4) and Bauer et al.\ (2016) for surveys.
}  
However, the lack of international recognition is one of the likely consequences from those civil wars in which one side attempts to secede from the parent country. 
By filling this gap in the literature, our study makes two contributions.

The literature finds that output loss due to conflicts tends to be temporary (Davis and Weinstein 2002; Miguel and Roland 2011; Cerra and Saxena 2008).%
\footnote{
	Mueller (2012), commenting on the methodology in Cerra and Saxena (2008), shows that output persistently drops after the \textit{beginning} of civil conflicts by over 15\%.
} 
These findings, however, may suffer from sample selection bias, as pointed out by Blattman and Miguel (2010): the data is unavailable for those countries that severely suffer from civil conflicts.   
Unrecognized states are an example of the out-of-the-sample cases of post-conflict situations. 
Consequently, understanding the consequences of unrecognized states will help us paint the full picture of the impacts of civil conflicts.

The literature on civil conflicts also pays little attention to the role of political institutions in post-conflict recovery.%
\footnote{
	An exception is Casey et al.\ (2012), who look at the impact of institutional building in post-conflict Sierra Leone.
}
Unrecognized states are a form of political institutions relevant for post-conflict periods. 
The present study thus contributes to the deepening of our understanding of the role of political institutions in post-conflict recovery.

The third literature that our study relates to is the political science literature on unrecognized states (Pegg 1998, King 2001, Lunch 2002, Kolsto 2006, Caspersen 2012, O'Loughlin et al.\ 2014, Smolnik 2016, among others). 
In particular, a recent strand of this literature points out that seeking international recognition creates an incentive for politicians in unrecognized states to promote state-building and democratic institutions (Caspersen 2008, 2011; Blakkisrud and Kolsto 2011; Kolsto and Blakkisrud 2012).
Our study complements this literature by uncovering economic consequences of non-recognition which can be interpreted as an indirect impact via state-building and democracy.%
\footnote{
	Witmer and O'Loughlin (2011) also use nighttime light data to gauge the impact of wars in towns in Georgia and South Ossetia. Our study differs in the coverage of study areas and in the recovery of economic growth rates from light data. 
}

@@

The rest of the paper is organized as follows. Section \ref{background} overviews the issue of unrecognized states in the former Soviet Union. Section \ref{theory} provides theoretical arguments on how non-recognition as a sovereign state can affect economic activities. Sections \ref{data} and \ref{methodology} introduce the datasets and methodology that we employ to uncover economic growth rates in unrecognized states. Section \ref{results} reports our findings. The last section concludes with discussions on the applicability of our findings to other unrecognized states.

>>>>

\begin{abstract}
This paper uses the satellite images of nighttime light to estimate economic growth rates in four unrecognized states of the former Soviet Union: Nagorno-Karabakh in Azerbaijan, Abkhazia and South Ossetia in Georgia, and Transnistria in Moldova. We then compare these estimates against those similarly obtained for the parent states to gauge the impact of non-recognition as sovereign states on economic activities. The estimated economic growth rates do not differ much between the breakaway territories and their parent states, suggesting that the economic impact of non-recognition as states may be fairly limited.
\end{abstract}

@@

\section{Introduction}
% Introducing frozen conflicts
Back in 1991, the Soviet Union broke up into fifteen of its constituent republics. 
Immediately after, wars of secession broke out in three of these newly independent states: Azerbaijan, Georgia, and Moldova.
Secessionists, consisting of minority ethnic groups in each republic, succeeded in taking over their territories and declared their independence from the parent states. 
However, not a single country around the world recognized them as sovereign states. 
The breakaway territories---Nagorno-Karabakh in Azerbaijan, Abkhazia and South Ossetia in Georgia, and Transnistria in Moldova (see Figure \ref{map})---remain internationally unrecognized as states until today.%
\footnote{%
	Since 2008, Russia and a few other countries have recognized Abkhazia and South Ossetia as sovereign states, though.
}

@@

The empirical challenge for such endeavor is the lack of data. International organizations such as the World Bank do not provide any data for unrecognized states. 
The limited state capacity of these territories implies that there is no official statistics. 
Any surveys conducted in their parent countries exclude these territories from the sampling procedure.%
\footnote{%
	As an example, the Demographic and Health Survey conducted in Moldova in 2005 excludes the entire Transnistria from its sample (National Scientific and Applied Center for Preventive Medicine 2006:\ xxvi).
} 
To overcome this data limitation, the present study uses the satellite images of nighttime light for the period 1992--2013 to recover the unobserved annual economic growth rates of the four unrecognized states in the former Soviet Union.

Specifically, we first estimate the elasticity of real GDP with respect to nighttime light intensity, conditional on country and year fixed effects, from the annual panel data of 177 countries around the world. 
Using this elasticity estimate and the estimated year fixed effects, we then recover the annual growth rates of real GDP for unrecognized states from their nighttime light intensity. 
We finally compare these recovered real GDP growth rates to the ones similarly estimated for the parent countries. 
Under the assumption that parent countries provide a valid counterfactual, the difference in the annual growth rates can be interpreted as the impact of being unrecognized as sovereign states.

We find that the impact of unrecognized states is fairly limited in all of the four cases. 
The recovered annual economic growth rates do not differ much between the breakaway territories and their parent states.

@@

Second, the issue of unrecognized states has received no attention in the emerging literature on the impacts of conflicts.%
\footnote{
	See Blattman and Miguel (2010:\ Section 4) and Bauer et al.\ (2016) for surveys.
}  
However, the lack of international recognition is one of the likely consequences from those civil wars in which a territory attempts to secede from its parent country. 
By filling this gap in the literature, our study makes two contributions.

The literature finds that output loss due to conflicts tends to be temporary (Davis and Weinstein 2002; Miguel and Roland 2011; Cerra and Saxena 2008).%
\footnote{
	Mueller (2012), commenting on the methodology in Cerra and Saxena (2008), shows that output persistently drops after the \textit{beginning} of civil conflicts by over 15\%.
} 
These findings, however, may suffer from sample selection bias, as pointed out by Blattman and Miguel (2010): the data is unavailable for those countries that severely suffer from civil conflicts.   
Unrecognized states are an example of the out-of-the-sample cases of post-conflict situations. 
Consequently, understanding the consequences of unrecognized states will help us paint the full picture of the impacts of conflicts.

The literature on civil conflicts also pays little attention to the role of political institutions in post-conflict recovery.%
\footnote{
	An exception is Casey et al.\ (2012), who look at the impact of institutional building in post-conflict Sierra Leone.
}
Unrecognized states are a form of political institutions relevant for the period after a ceasefire. 
The present study thus contributes to the deepening of our understanding of the role of political institutions in post-conflict recovery.

The third literature that our study relates to is the political science literature on unrecognized states (Pegg 1998, King 2001, Lynch 2002, Kolsto 2006, Caspersen 2012, O'Loughlin et al.\ 2014, Smolnik 2016, among others). 
In particular, a recent strand of this literature points out that seeking international recognition creates an incentive for politicians in unrecognized states to promote state-building and democratic institutions (Caspersen 2008, 2011; Blakkisrud and Kolsto 2011; Kolsto and Blakkisrud 2012).
Our study complements this literature by uncovering economic consequences of non-recognition which can be interpreted as an indirect impact via state-building and democracy.%
\footnote{
	Witmer and O'Loughlin (2011) also use nighttime light data to gauge the impact of wars in towns in Georgia and South Ossetia. Our study differs in the coverage of study areas and in the recovery of economic growth rates from light data. 
}

@@

The rest of the paper is organized as follows. Section \ref{background} overviews the issue of unrecognized states in the former Soviet Union. Section \ref{theory} provides theoretical arguments on how non-recognition as a sovereign state can affect economic activities. Sections \ref{data} and \ref{methodology} introduce the datasets and methodology, respectively, that we employ to uncover economic growth rates in unrecognized states. Section \ref{results} reports our findings, followed by the concluding section.
